---
title: "Assessment02_pt_b"
author: "Xilin Huang Liam"
date: "20/09/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data from Yahoo Finance
Before we starting working on building models and analysis, data need to be loaded from Yahoo Finance API. In our project, we will be collecting:

Coins (to US dollar)

* BTC-USD: Bitcoin

* ETH-USD: Ethereum

* ADA-USD: Cardano

* DOGE-USD: Dogecoin

* SHIB-USD: Shiba Inu coin

Index

* ^DJI: Dow Jones Industrial

* ^IXIC: Nasdaq Composite

* ^GSPC: S&P 500

* GC=F: Gold

Stocks

* TSLA: Tesla

* GOOG: Google

* AAPL: Apple

* NVDA: Nvidia

* AMD: Advanced Micro Devices

* TSM: Taiwan Semiconductor Manufacturing

All the data will be saved as `xts` (Extensible Time Series) object.

```{r yahoo, message=FALSE, warning=FALSE}
rm(list=ls())
library(zoo)
library(xts)
library(TTR)
library(quantmod)

coin_portfolio=c("BTC-USD","ETH-USD","ADA-USD",
                 "DOGE-USD","SHIB-USD")
index_portfolio=c("^DJI","^IXIC","^GSPC","GC=F")
stock_portfolio=c("TSLA","GOOG","AAPL","NVDA","AMD","TSM")

data <- getSymbols(c(coin_portfolio,
                     index_portfolio,
                     stock_portfolio),
                   src='yahoo',
                   #from=dyear,
                   #to=d,
                   autoassign=FALSE)
```

## Simple linear regression in finance data
For simple linear regression, we will use TSLA - Bitcoin as example. Since this project will be focusing on performing linear regression models, the data will be transformed from `xts` object to `dataframe`. Other data will be transformed to dataframes for multi-linear model analysis based on the result of TSLA - Bitcoin.

```{r data}
df_BTC = data.frame(date=index(`BTC-USD`), coredata(`BTC-USD`))
df_TSLA = data.frame(date=index(TSLA), coredata(TSLA))
# coins
df_ETH = data.frame(date=index(`ETH-USD`), coredata(`ETH-USD`))
df_ADA = data.frame(date=index(`ADA-USD`), coredata(`ADA-USD`))
df_DOGE = data.frame(date=index(`DOGE-USD`), coredata(`DOGE-USD`))
df_SHIB = data.frame(date=index(`SHIB-USD`), coredata(`SHIB-USD`))
# indexs
df_DJI = data.frame(date=index(DJI), coredata(DJI))
df_IXIC = data.frame(date=index(IXIC), coredata(IXIC))
df_GSPC = data.frame(date=index(GSPC), coredata(GSPC))
df_GOLD = data.frame(date=index(`GC=F`), coredata(`GC=F`))
# stock
df_GOOG = data.frame(date=index(GOOG), coredata(GOOG))
df_AAPL = data.frame(date=index(AAPL), coredata(AAPL))
df_NVDA = data.frame(date=index(NVDA), coredata(NVDA))
df_AMD = data.frame(date=index(AMD), coredata(AMD))
df_TSM = data.frame(date=index(TSM), coredata(TSM))
```

```{r friday_filled, message=FALSE, warning=FALSE}
library(tidyr)
# use Friday's data for weekends
df_BTCTSLA <- merge(df_BTC, df_TSLA, by = 'date', all.x = TRUE)
df_BTCTSLA_filled <- df_BTCTSLA %>%
fill(TSLA.Open,TSLA.High,TSLA.Low,TSLA.Close,TSLA.Adjusted,TSLA.Volume)
```

```{r 2019, message=FALSE, warning=FALSE}
# subset data (2019, close price and volume)
df_BTCTSLA_sub <- subset(df_BTCTSLA_filled,
                         date >= '2019-01-01',
                         select = c(date,
                                    TSLA.Close,
                                    TSLA.Volume,
                                    BTC.USD.Close,
                                    BTC.USD.Volume))
row.names(df_BTCTSLA_sub) <- NULL
```

```{r lag_data, message=FALSE, warning=FALSE}
library(dplyr)
lag_list = c(1, 3, 5, 10, 20, 30, 50, 60, 90, 100)
for (i in lag_list){
  if (i == lag_list[1]) {
    df_BTCTSLA_lag = data.frame(col1 = lag(df_BTCTSLA_sub$TSLA.Close, n = i))
    names(df_BTCTSLA_lag)[ncol(df_BTCTSLA_lag)] <- paste0("TSLA_price_lag_", i)
  } else {
    df_BTCTSLA_lag[,ncol(df_BTCTSLA_lag)+1] <- lag(df_BTCTSLA_sub$TSLA.Close, n = i)
    names(df_BTCTSLA_lag)[ncol(df_BTCTSLA_lag)] <- paste0("TSLA_price_lag_", i)
  }
}
```

```{r lag_data_merge, echo=TRUE}
df_BTCTSLA_lag_m <- cbind(df_BTCTSLA_sub, df_BTCTSLA_lag)
```

```{r return_data, echo=TRUE, message=FALSE}
# Bitcoin daily return
BDReturn <- c(NA, diff(df_BTCTSLA_lag_m$BTC.USD.Close)/df_BTCTSLA_lag_m$BTC.USD.Close*100)
BDReturn <- BDReturn[-length(BDReturn)]
df_BTCTSLA_lag_m$BTC_USD_Day_Return <- BDReturn
# TSLA daily return
TDReturn <- c(NA, diff(df_BTCTSLA_lag_m$TSLA.Close)/df_BTCTSLA_lag_m$TSLA.Close*100)
TDReturn <- TDReturn[-length(TDReturn)]
df_BTCTSLA_lag_m$TSLA_price_Day_Return <- TDReturn
```

```{r lm_model, echo=TRUE, warning=FALSE}
library(ggplot2)
library(ggfortify)
# Best result at 90 lag days -- log(y) ~ x
x_value1 = df_BTCTSLA_lag_m$TSLA_price_lag_90
y_value1 = log(df_BTCTSLA_lag_m$BTC.USD.Close)
fit1 = lm(data = df_BTCTSLA_lag_m, y_value1 ~ x_value1)
summary(fit1)
ggplot(df_BTCTSLA_lag_m, aes(x = x_value1, y = y_value1)) +
  geom_point() +
  stat_smooth(method = 'lm', col = 'red') + 
  labs(title = paste("Adj R2 = ", signif(summary(fit1)$adj.r.squared, 5),
                     " Intercept = ", signif(fit1$coef[[1]], 5),
                     " Slope =", signif(fit1$coef[[2]], 5),
                     " P =", signif(summary(fit1)$coef[2,4], 5)))
# Relative Standard Error
sigma(fit1)/mean(df_BTCTSLA_lag_m$BTC.USD.Close, na.rm=TRUE)
autoplot(fit1)
```

```{r lm_model_log, echo=TRUE, warning=FALSE}
# Best result at 60 days -- y ~ x
x_value2 = df_BTCTSLA_lag_m$TSLA_price_lag_60
y_value2 = df_BTCTSLA_lag_m$BTC.USD.Close
fit2 = lm(data = df_BTCTSLA_lag_m, y_value2 ~ x_value2)
summary(fit2)
ggplot(df_BTCTSLA_lag_m, aes(x = x_value2, y = y_value2)) +
  geom_point() +
  stat_smooth(method = 'lm', col = 'red') + 
  labs(title = paste("Adj R2 = ", signif(summary(fit2)$adj.r.squared, 5),
                     " Intercept = ", signif(fit2$coef[[1]], 5),
                     " Slope =", signif(fit2$coef[[2]], 5),
                     " P =", signif(summary(fit2)$coef[2,4], 5)))
# Relative Standard Error
sigma(fit2)/mean(df_BTCTSLA_lag_m$BTC.USD.Close, na.rm=TRUE)
autoplot(fit2)
```

```{r lm_model_return, echo=TRUE, warning=FALSE}
x_value3 = df_BTCTSLA_lag_m$TSLA_price_Day_Return
y_value3 = df_BTCTSLA_lag_m$BTC_USD_Day_Return
fit3 = lm(data = df_BTCTSLA_lag_m, y_value3 ~ x_value3)
summary(fit3)
ggplot(df_BTCTSLA_lag_m, aes(x = x_value3, y = y_value3)) +
  geom_point() +
  stat_smooth(method = 'lm', col = 'red') + 
  labs(title = paste("Adj R2 = ", signif(summary(fit3)$adj.r.squared, 5),
                     " Intercept = ", signif(fit3$coef[[1]], 5),
                     " Slope =", signif(fit3$coef[[2]], 5),
                     " P =", signif(summary(fit3)$coef[2,4], 5)))
# Relative Standard Error
sigma(fit3)/mean(df_BTCTSLA_lag_m$BTC_USD_Day_Return, na.rm=TRUE)
autoplot(fit3)
```

```{r others}
# coins data frame
df_coin = data.frame()
df_coin = merge(df_BTC, df_ETH, by = 'date', all.x = TRUE)
df_coin = merge(df_coin, df_ADA, by = 'date', all.x = TRUE)
df_coin = merge(df_coin, df_DOGE, by = 'date', all.x = TRUE)
df_coin = merge(df_coin, df_SHIB, by = 'date', all.x = TRUE)
# indexes data frame
df_index = data.frame()
df_index = merge(df_DJI, df_IXIC, by = 'date', all.x = TRUE)
df_index = merge(df_index, df_GSPC, by = 'date', all.x = TRUE)
df_index = merge(df_index, df_GOLD, by = 'date', all.x = TRUE)
# stocks data frame
df_stock = data.frame()
df_stock = merge(df_TSLA, df_GOOG, by = 'date', all.x = TRUE)
df_stock = merge(df_stock, df_AAPL, by = 'date', all.x = TRUE)
df_stock = merge(df_stock, df_NVDA, by = 'date', all.x = TRUE)
df_stock = merge(df_stock, df_AMD, by = 'date', all.x = TRUE)
df_stock = merge(df_stock, df_TSM, by = 'date', all.x = TRUE)
# merge 3 portfolios data frame
df_port = data.frame()
df_port = merge(df_coin, df_index, by = 'date', all.x = TRUE)
df_port = merge(df_port, df_stock, by = 'date', all.x = TRUE)

# subset 2019
df_port <- subset(df_port,
                  date >= '2019-01-01',
                  select = c(date, 
                             BTC.USD.Close, 
                             BTC.USD.Volume,
                             ETH.USD.Close, 
                             ETH.USD.Volume,
                             ADA.USD.Close, 
                             ADA.USD.Volume,
                             DOGE.USD.Close, 
                             DOGE.USD.Volume,
                             SHIB.USD.Close, 
                             SHIB.USD.Volume,
                             DJI.Close, 
                             DJI.Volume,
                             IXIC.Close, 
                             IXIC.Volume,
                             GSPC.Close, 
                             GSPC.Volume,
                             GC.F.Close, 
                             GC.F.Volume,
                             TSLA.Close, 
                             TSLA.Volume,
                             GOOG.Close, 
                             GOOG.Volume,
                             AAPL.Close, 
                             AAPL.Volume,
                             NVDA.Close, 
                             NVDA.Volume,
                             AMD.Close, 
                             AMD.Volume,
                             TSM.Close, 
                             TSM.Volume
                             ))
row.names(df_port) <- NULL
```

```{r multi_linear}
m_fit <- lm(`BTC.USD.Close` ~ . -date,
            data = df_port)
summary(m_fit)
```

```{r multi_linear_stars}
m_fit_adjusted <- lm(`BTC.USD.Close` ~ 
                       `BTC.USD.Volume` + 
                       `ETH.USD.Close` + 
                       `ETH.USD.Volume` + 
                       `DOGE.USD.Close` +
                       `GC.F.Close`,
                     data = df_port)
summary(m_fit_adjusted)
# Relative Standard Error
sigma(m_fit_adjusted)/mean(df_port$BTC.USD.Close, na.rm=TRUE)
```
```{r multi_linear_stars_f}
m_fit_adjusted_filtered <- lm(`BTC.USD.Close` ~ 
                                `BTC.USD.Volume` +
                                `ETH.USD.Close` + 
                                `DOGE.USD.Close`,
                              data = df_port)
summary(m_fit_adjusted_filtered)
# Relative Standard Error
sigma(m_fit_adjusted_filtered)/mean(df_port$BTC.USD.Close, na.rm=TRUE)
```